The Implementation of China’s Overseas NGO Law and the Operating Space for International Civil Society

### 2023-08-23 manuscript/manuscript.qmd
updated manuscript

Many INGOs gained legal identity within mainland China using the vehicle of foreign-funded companies [@Sidel:2016], though there were no clear requirements for them to have a formal legal presence in China [@Han:2011]. However, the legal form of foreign-funded company is not quite compatible with the nonprofit nature of INGOs. Under the Chinese legal system, enterprises are intended to be for-profit entities and nonprofit organizations are required to choose other stand-alone legal forms, such as foundations, associations, and social service organizations [@Ye:2021]. Thus, INGOs obtaining legal identity in this way have to navigate through a legal framework designed for a for-profit institutional logic, e.g, focusing on project investors' right to investment returns.

The legal framework governing INGOs was overhauled by the 2017 Chinese ONGO Law. The new law applies to all overseas (including Hong Kong, Macau, and Taiwan) non-governmental organizations that conduct not-for-profit activities in the fields of economic development, education, science, technology, culture, public health, sports, environment protection, poverty alleviation, and disaster relief (Article 3). The ONGO Law stipulates two ways for INGOs to lawfully conduct nonprofit activities in China: (1) submitting an application for formal permanent registration, or (2) filing documentation of temporary activities with the Ministry of Public Security and provincial-level Public Security Bureaus, which are the registering and regulatory authority for INGOs. Formal registration an RO with a legal identity and allows them to th maintain an ongoing present in the country, while temporary activities are suitable for INGOs that need to conduct one-off programs that last no longer than one year. The law's registration requirements were strictly enforced—after the Overseas NGO Law came into effect, any INGOs that could not successfully obtain proper legal authorization would be deemed illegal.

This paper explores the characteristics of all registered INGOs in China in the five years after the ONGO Law became effective and examings the association between INGO issue areas and the operational space granted to them in the country. Registering as an RO under the ONGO Law is a complex process that requires the consent of a Chinese Professional Supervisory Unit (PSU)—a governmental or quasi-governmental unit related to an INGO's issue area which oversees and co-supervises the INGO with Public Security Bureaus. Obtaining the consent of a PSU to apply for registration has become a key challenge for INGOs interested in formal registration, since potential PSUs do not have a legal obligation to assume such a demanding role [@Jia:2017; @Shieh:2018; @YeHuang:2018]. INGOs vary in their ability to attract partner PSUs, and organizations that are unable to find corresponding PSUs are more limited in their geographic reach within the country. INGOs are authorized to operate within China on a province-by-province basis. Many organizations are allowed to work in all 32 mainland provincial administrative regions, but many are also only authorized to work in a handful, or even only a single province. We use the number of provinces that INGOs are authorized to work in as a proxy to measure the operational space for INGOs in China based on the assumption that INGOs need to or hope to have the approval to be able to carry out their programs in a geographical scope as large as possible [@Li:2020].

@@

Based on extant literature and theoretical understanding, we argue that the operation space INGOs are granted is associated with the level of contentious their missions, i.e., the issue areas they work in. Since the level of contentiousness is also manifested by the level of difficulty they face while seeking lawful registration.

@@

The data used in this paper was collected from INGOs' public registration documents disclosed by the Ministry of Public Security's Overseas NGO Office online service platform. The data spans from January 1, 2017, when the law took effect, to December 31, 2021, with data for 633 registered organizations. The filing documents for ROs disclose the name, the address, the registration date, the registration authority (which public security bureau regulates it), the PSU, the issue area or purposes of the RO, and the specific provinces they can work in. We merge our INGO data with additional data from ChinaFile, including English organization names, for the sake of transparency and accessibility.

We code the variables we use in the empirical analysis from the registration information. Most of our data cleaning work is an objective data wrangling process, such as parsing out the province each RO is in from the address, or calculating the total number provinces approved in the registration document. Meanwhile, the qualitative coding of issue areas is conducted by the authors

@@

Our second variable is a dichotomous indicator of whether the INGO explicitly states that it intends to operate locally. There are cases where the aim or mission of the INGO attached to a locality, either having a county or city name in their RO name or purposes, or the INGO are actually set up by people in China and they registered the RO to communicate between the "hometown" and the INGO set up abroad. We thus include this variable in our analysis to explore how having local connections is associated with the operational space of ROs.

Our final variable measures how much time elapsed between the January 1, 2017 (when the law took effect) and the organization's official registration date. This indicates how long it took INGOs to seek the consent and supervisory authority of both a PSU and the Public Security Bureau and to complete all other necessary bureaucratic requirements.

>>>>

Many INGOs gained legal identity within mainland China using the vehicle of foreign-funded companies [@Sidel:2016], though there were no clear requirements for them to have a formal legal presence in China [@Han:2011]. However, the legal form of foreign-funded company is not quite compatible with the nonprofit nature of INGOs. Under the Chinese legal system, enterprises are intended to be for-profit entities and nonprofit organizations are required to choose other stand-alone legal forms, including foundations, associations, and social service organizations [@Ye:2021]. Thus, INGOs obtaining legal identity in this way have to navigate through a legal framework designed for a for-profit institutional logic, e.g, focusing on protecting investors' right to investment returns.

The legal framework governing INGOs was overhauled by the 2017 Chinese ONGO Law. The new law applies to all overseas (including Hong Kong, Macau, and Taiwan) non-governmental organizations that conduct not-for-profit activities in the fields of economic development, education, science, technology, culture, public health, sports, environment protection, poverty alleviation, and disaster relief (Article 3). The ONGO Law stipulates two ways for INGOs to lawfully conduct nonprofit activities in China: (1) submitting an application for formal permanent registration, or (2) filing documentation of temporary activities with the Ministry of Public Security and provincial-level Public Security Bureaus, which are the registering and regulatory authority for INGOs. Formal registration of an RO allows them to maintain an ongoing present in the country, while temporary activities are intended for INGOs that need to conduct one-off programs that last no longer than one year. Abiding by these two channels to obtain lawful presence is mandatory for INGOs after the Overseas NGO Law came into effect, any INGOs that could not successfully obtain proper legal authorization would be deemed illegal.

This paper explores the characteristics of all registered INGOs in China in the five years after the ONGO Law became effective and examines the association between INGO issue areas and the operational space granted to them in the country. Registering as an RO under the ONGO Law is a complex process that requires the consent of a Chinese Professional Supervisory Unit (PSU)—a governmental or quasi-governmental unit related to an INGO's issue area which oversees and co-supervises the INGO with Public Security Bureaus. Obtaining the consent of a PSU to apply for registration has become a key challenge for INGOs interested in formal registration, since potential PSUs do not have a legal obligation to assume such a demanding role [@Jia:2017; @Shieh:2018; @YeHuang:2018]. INGOs vary in their ability to attract partner PSUs, and organizations that are unable to find corresponding PSUs are more limited in their geographic reach within the country. INGOs are authorized to operate within China on a province-by-province basis. Many organizations are allowed to work in all 32 mainland provincial administrative regions, but many are also only authorized to work in a handful, or even only a single province. We use the number of provinces that INGOs are authorized to work in as a proxy to measure the operational space for INGOs in China based on the assumption that INGOs need to or hope to have the approval to be able to carry out their programs in a geographical scope as large as possible [@Li:2020].

@@

Based on extant literature and theoretical understanding, we argue that the operation space INGOs are granted is associated with the level of contentiousness their missions, i.e., the issue areas they work in. Since the level of contentiousness is also manifested by the level of difficulty they face while seeking lawful registration.

@@

The data used in this paper was collected from INGOs' public registration documents disclosed by the Ministry of Public Security's Overseas NGO Office online service platform. The platform is a website maintained by the registrar of INGOs and regularly publicize the administrative documentation of all the ROs and temporary activities registered or filled across China. The data used in this study spans from January 1, 2017, when the law took effect, to December 31, 2021, with information for all 633 registered organizations approved as of the cutoff point. The filing documents for ROs disclose the name, the address, the registration date, the registration authority (which public security bureau regulates it), the PSU, the issue area or purposes of the RO, and the specific provinces they can work in. We merge our INGO data with additional data of Chinese INGO data from ChinaFile, a project that translate and code the MPS platform data, to include their English organization names, for the sake of transparency and accessibility.

We code the variables we use in the empirical analysis from the registration information. Most of our data cleaning work is an objective data wrangling process, such as parsing out the province each RO registered in from the address, or calculating the total number provinces approved in the registration document. Meanwhile, the qualitative coding of issue areas is conducted by the authors, based on consistent coding rules agreed by the authors.

@@

Our second independent variable is a dichotomous indicator of whether the INGO explicitly states that it intends to operate locally (below province level). There are cases where the aim of the INGO is attached to a locality, either having a county or city name in their RO name or purposes, or the INGO are actually set up by people in China and they registered the RO to communicate between the "hometown" and the INGO set up abroad. We thus include this variable in our analysis to explore how having local connections is associated with the operational space of ROs.

Our final independent variable measures how much time elapsed between the January 1, 2017 (when the law took effect) and the organization's official registration date. There is great variance in the timing of registering an RO among INGOs, since there has been a *de facto* grace period for INGOs to seek legal presence after the law took effect. This is largely because both INGOs and the regulators needed some transitional time to explore the specific implementation of the law as large ambiguity was left when the law was passed. We use this temporal variable as a proxy for the motivation or eagerness of INGOs to register an RO, for some INGOs are inclined to avoid the cumbersome registration procedures at the early stage if there is no real consequences for continuing operations without registration, while others took the time to talk with multiple potential PSUs to find a more suitable superior or regulatory dynamic. (to find a citation for this, probably Spires, if there is one.)
